Effectiveness of visual test in linear regression diagnositics

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.Rmd
fin data preprocessing

As pointed out by @vanderplas2021statistical, though equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. They summarized three common different scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design. For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, they decided to add $1/s_j$ to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$.

@@

Note that the use of equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features. They suggested to estimate $\hat{\alpha}$ visually based on the selections of null plots of the experimental data, or to estimate $\hat{\alpha}$ numerically based on several additional Rorschach lineups, which is a type of lineup containing only null plots. However, when the number of null models are large, it could be expensive to manually estimate each $\alpha$ or include additional Rorschach lineups in the experiment.

@@

As suggested by many previous studies [many studies refs here], it is reasonable for an subject to evaluate a block of 20 lineups. Subjects for both two experiments were recruited from Prolific web service [ref here]. They were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others. If there was no noticeable difference between plots in a lineup, subject could choose to select zero lineups without providing the reason. Information about gender, age, education, and previous experience in visual experiment were also collected. No subject was shown the same lineup twice.

@@

Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups. For each lineup, the actual data plot was drawn as standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot. Figure \ref{fig:example-lineup} is an example of one of these lineups. It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $N(0.0.3^2)$ as the distribution of $X_{raw}$. The actual data plot location is four. All five subjects correctly identified the actual data plot for this lineup.

@@

Subjects recruited from Prolific received a fixed payment for submitting quality responses. However, some subjects will try to maximize their earnings for minimum effort. If a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed both attention checks, we rejected the submission and the subject would not be paid.

Participants recruited from Prolific are paid for their efforts. Some participants will try to maximize their earnings for minimum effort, which can affect the result from the data. For the purpose of identifying these participants and cleaning the data, two attention checks are included in the block of 20 lineups. If the subject failed to identify at least one of the actual data plot in those attention checks, we reject the submission and the subject will not be paid. All attention checks are removed from the final data. [table ref here] tabulates the number of subjects, genders, and lineups evaluated after applying the data screening procedure.

@@

>>>>

As pointed out by @vanderplas2021statistical, though Equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. They summarized three common different scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design. For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, they decided to add $1/s_j$ to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$.

@@

Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features. They suggested to estimate $\hat{\alpha}$ visually based on the selections of null plots of the experimental data, or to estimate $\hat{\alpha}$ numerically based on several additional Rorschach lineups, which is a type of lineup containing only null plots. However, when the number of null models are large, it could be expensive to manually estimate each $\alpha$ or include additional Rorschach lineups in the experiment.

@@

As suggested by many previous studies [many studies refs here], it is reasonable for an subject to evaluate a block of 20 lineups. Subjects for both two experiments were recruited from Prolific web service [ref here]. Data screening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions. During the experiment, they were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others. If there was no noticeable difference between plots in a lineup, subject could choose to select zero plots without providing the reason. Information about gender, age, education, and previous experience in visual experiment were also collected. No subject was shown the same lineup twice.

@@

Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups. For each lineup, the actual data plot was drawn as standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot. Figure \ref{fig:example-lineup} is an example of one of these lineups. It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location is four. All five subjects correctly identified the actual data plot for this lineup.

@@

Subjects recruited from Prolific received a fixed payment for participating in the experiment. However, some subjects will try to maximize their earnings for minimum effort. During the review of submissions, if we found a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed all attention checks, we rejected the submission. The rejected submissions will be removed immediately, and Prolific will automatically recruit another subject as substitution. Therefore, we only paid for approved submissions and no further data screening procedure needed to be applied on the collected data.

A subject was allowed to select zero plots for a lineup, which literally means there was no visble difference between plots,
but the simulated p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop the subject from the simulation for this case, leading to inaccurate estimation of $p$-value. Therefore, we treated this case as making one selection but failing to identify the actual data plot so that Equation (\ref{eq:p-value-multi}) can be applied correctly.

@@

[table ref here] tabulates the number of subjects, genders, and lineups evaluated after applying the data screening procedure.

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.tex
fin data preprocessing

As pointed out by \citet{vanderplas2021statistical}, though equation
(\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the
possible dependencies in the visual test due to repeated evaluations of
the same lineup. And it is inapplicable to visual test where subjects
are asked to select one or more ``most different'' plots from the
lineup. They summarized three common different scenarios in visual
inference: (1) \(K\) different lineups are shown to \(K\) subjects, (2)
\(K\) lineups with different null plots but the same actual data plot
are shown to \(K\) subjects, and (3) the same lineup is shown to \(K\)
subjects. Out of these three scenarios, Scenario 3 is the most common in
previous studies as it puts the least constraints on the experimental
design. For Scenario 3, \citet{vanderplas2021statistical} modelled the
probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, they decided to add \(1/s_j\) to
\(c_i\) instead of one, where \(s_j\) is the number of plots subject
\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\).

@@

Note that the use of equation (\ref{eq:pvalue-beta-binomial}) requires
the estimation of \(\hat{\alpha}\), which largely depends on the null
model, the type of the plot and other aesthetic features. They suggested
to estimate \(\hat{\alpha}\) visually based on the selections of null
plots of the experimental data, or to estimate \(\hat{\alpha}\)
numerically based on several additional Rorschach lineups, which is a
type of lineup containing only null plots. However, when the number of
null models are large, it could be expensive to manually estimate each
\(\alpha\) or include additional Rorschach lineups in the experiment.

@@

As suggested by many previous studies {[}many studies refs here{]}, it
is reasonable for an subject to evaluate a block of 20 lineups. Subjects
for both two experiments were recruited from Prolific web service {[}ref
here{]}. They were asked to select one or more plots that are mostly
different from others, provide a reason for their selections, and
express how different they think the selected plots are from others. If
there was no noticeable difference between plots in a lineup, subject
could choose to select zero lineups without providing the reason.
Information about gender, age, education, and previous experience in
visual experiment were also collected. No subject was shown the same
lineup twice.

@@

Three replications are made for each of the parameter values shown in
Table \ref{tab:parameter-table} resulting in 192 different lineups. For
each lineup, the actual data plot was drawn as standard residual plot of
the null model with raw residuals on the y-axis and fitted values on the
x-axis. The 19 null datasets were generated by the residual rotation
technique, and plotted in the same way. The lineup consisted of 20
residual plots with one randomly placed actual data plot. Figure
\ref{fig:example-lineup} is an example of one of these lineups. It was
produced by using \(n = 300\), \(j = 6\), \(\sigma = 0.5\) and
\(N(0.0.3^2)\) as the distribution of \(X_{raw}\). The actual data plot
location is four. All five subjects correctly identified the actual data
plot for this lineup.

@@

Subjects recruited from Prolific received a fixed payment for submitting
quality responses. However, some subjects will try to maximize their
earnings for minimum effort. If a subject objectively demonstrated clear
low-effort throughout the experiment, i.e., failed both attention
checks, we rejected the submission and the subject would not be paid.

Participants recruited from Prolific are paid for their efforts. Some
participants will try to maximize their earnings for minimum effort,
which can affect the result from the data. For the purpose of
identifying these participants and cleaning the data, two attention
checks are included in the block of 20 lineups. If the subject failed to
identify at least one of the actual data plot in those attention checks,
we reject the submission and the subject will not be paid. All attention
checks are removed from the final data. {[}table ref here{]} tabulates
the number of subjects, genders, and lineups evaluated after applying
the data screening procedure.

@@

>>>>

As pointed out by \citet{vanderplas2021statistical}, though Equation
(\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the
possible dependencies in the visual test due to repeated evaluations of
the same lineup. And it is inapplicable to visual test where subjects
are asked to select one or more ``most different'' plots from the
lineup. They summarized three common different scenarios in visual
inference: (1) \(K\) different lineups are shown to \(K\) subjects, (2)
\(K\) lineups with different null plots but the same actual data plot
are shown to \(K\) subjects, and (3) the same lineup is shown to \(K\)
subjects. Out of these three scenarios, Scenario 3 is the most common in
previous studies as it puts the least constraints on the experimental
design. For Scenario 3, \citet{vanderplas2021statistical} modelled the
probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, they decided to add \(1/s_j\) to
\(c_i\) instead of one, where \(s_j\) is the number of plots subject
\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\).

@@

Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires
the estimation of \(\hat{\alpha}\), which largely depends on the null
model, the type of the plot and other aesthetic features. They suggested
to estimate \(\hat{\alpha}\) visually based on the selections of null
plots of the experimental data, or to estimate \(\hat{\alpha}\)
numerically based on several additional Rorschach lineups, which is a
type of lineup containing only null plots. However, when the number of
null models are large, it could be expensive to manually estimate each
\(\alpha\) or include additional Rorschach lineups in the experiment.

@@

As suggested by many previous studies {[}many studies refs here{]}, it
is reasonable for an subject to evaluate a block of 20 lineups. Subjects
for both two experiments were recruited from Prolific web service {[}ref
here{]}. Data screening procedure was applied during the recruitment,
subjects were required to be fluent in English, with \(98\%\) minimum
approval rate and 10 minimum submissions. During the experiment, they
were asked to select one or more plots that are mostly different from
others, provide a reason for their selections, and express how different
they think the selected plots are from others. If there was no
noticeable difference between plots in a lineup, subject could choose to
select zero plots without providing the reason. Information about
gender, age, education, and previous experience in visual experiment
were also collected. No subject was shown the same lineup twice.

@@

Three replications are made for each of the parameter values shown in
Table \ref{tab:parameter-table} resulting in 192 different lineups. For
each lineup, the actual data plot was drawn as standard residual plot of
the null model with raw residuals on the y-axis and fitted values on the
x-axis. The 19 null datasets were generated by the residual rotation
technique, and plotted in the same way. The lineup consisted of 20
residual plots with one randomly placed actual data plot. Figure
\ref{fig:example-lineup} is an example of one of these lineups. It was
produced by using \(n = 300\), \(j = 6\), \(\sigma = 0.5\) and
\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location is four. All
five subjects correctly identified the actual data plot for this lineup.

@@

Subjects recruited from Prolific received a fixed payment for
participating in the experiment. However, some subjects will try to
maximize their earnings for minimum effort. During the review of
submissions, if we found a subject objectively demonstrated clear
low-effort throughout the experiment, i.e., failed all attention checks,
we rejected the submission. The rejected submissions will be removed
immediately, and Prolific will automatically recruit another subject as
substitution. Therefore, we only paid for approved submissions and no
further data screening procedure needed to be applied on the collected
data.

A subject was allowed to select zero plots for a lineup, which literally
means there was no visble difference between plots, but the simulated
p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop
the subject from the simulation for this case, leading to inaccurate
estimation of \(p\)-value. Therefore, we treated this case as making one
selection but failing to identify the actual data plot so that Equation
(\ref{eq:p-value-multi}) can be applied correctly.

@@

{[}table ref here{]} tabulates the number of subjects, genders, and
lineups evaluated after applying the data screening procedure.
